Inferring the gravitational potential of the Milky Way with a few precisely measured stars

### 2014-02-13 text/paper2/ms.tex
affiliations

\title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston}

@@

The difficulty and computational costs associated with proper N-body modeling has motivated numerous efforts for developing approximate methods for constraining properties of the Galactic potential with cold debris structures. The simplest approximation is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}. This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{eyre11, sanders13a}. To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14}. Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).

@@

In \citet[][hereafter \paperone]{apw13}, we introduced a simple method for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential (\emph{Rewinder}). [...]

[We properly handle observational errors]

>>>>

\title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}, 
	    David W. Hogg\altaffilmark{\nyu,\mpia}, 
	    Kathryn V. Johnston\altaffilmark{\colum}, 
	    David Hendel\altaffilmark{\colum}}

% Affiliations
\newcommand{\colum}{1}
\newcommand{\adrn}{2}
\newcommand{\nyu}{3}
\newcommand{\mpia}{4}
\altaffiltext{\colum}{Department of Astronomy, 
		              Columbia University, 
		              550 W 120th St., 
		              New York, NY 10027, USA}
\altaffiltext{\adrn}{To whom correspondence should be addressed: \email{adrn@astro.columbia.edu}}
\altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
                      Department of Physics, New York University,
                      4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}

@@

The difficulty and computational costs associated with proper N-body modeling has motivated numerous efforts for developing approximate methods for constraining properties of the Galactic potential with cold debris structures. The simplest approximation is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}. This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{eyre11, sanders13a}. To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}. Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).

@@

In \citet[][hereafter \paperone]{apw13}, we introduced a simple method (\emph{Rewinder}) for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential. The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity). In this paper, we investigate how these relationships scale with the mass and orbit of the progenitor and the quality of the data, then present a new probabilistic framework for simultaneously modeling stream debris and the parent potential based on the same arguments.

In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses over two characteristic orbits. In Section~\ref{sec:}
